
```{r ,echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(corrplot)
library(gridExtra)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
merged <- read_csv("winequality_all.csv")
merged$red <- as.factor(merged$red )
fulldf <- merged %>% mutate(quality_group = cut(quality, breaks = c(0,4,6,10),labels=c("lo", "mid", "hi"))) 

```


```{R, echo = FALSE, message = FALSE}
data <- fulldf
data %>% 
  ggplot(aes(chlorides)) +
  geom_histogram(binwidth = 0.01, color = "blue",fill = "steelblue") 

data %>% 
  ggplot(aes(log2(chlorides))) +
  geom_histogram(binwidth = 0.5, color = "blue",fill = "steelblue") 
```

We see that the distribution of chloride is right skewed for white wine group, not normally distributed. After applying log2 transformation, the distribution of residual sugar is less right skewed and looks more like a normal distribution.

```{R, echo = FALSE, message = FALSE}
ggplot(data = data, aes(x = quality_group, y = log2(chlorides), color = quality_group)) +
  geom_boxplot()
```

Since there are plenty of outliers, we transform into log2 form and can see that low and middle quality group have similar means of chloride. The middle quality group has more outliers and the high quality group has a smaller level of chloride, on average. All the quality groups have similar variance of chloride.
