
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```



```{r}

require(factoextra)
require(NbClust)
require(cluster)
require(dplyr)
require(knitr)
require(ggplot2)
require(ggradar)


```

# Clustering Tendency

First, we use the Hopkins statistic (Lawson and Jurs 1990) denoted as H to assess the clustering tendency. If there is no clustering tendency, H would be about 0.5. A value for H higher than 0.75 indicates a clustering tendency at the 90% confidence level(Han, Kamber, and Pei 2012).

```{r}

data = read.csv("winequality_all.csv")
data = data[, -1]
data = data[, 1:(ncol(data) - 3)] # Remove last two columns
data = scale(data)# scaled the data
data_df = data.frame(data)
```

```{r}
library(factoextra)
res = get_clust_tendency(data, n = 100, graph = FALSE)
round(res$hopkins_stat,3)
```


# The Elbow Method

For the elbow method, in which the sum of squares at each number of clusters is calculated and graphed, we can look for a change of slope from steep to shallow to determine the optimal number of clusters. 

```{r}
library(factoextra)
fviz_nbclust(data, kmeans, method = "wss") + geom_vline(xintercept = 3, linetype = 2) + labs(subtitle = "Elbow method") 

```

The optimal cluster number given by the elbow method is 3.


# The Silhouette Method

The Silhouette method computes the average silhouette of observations for different numbers. The optimal number of clusters is the one that maximizes the average silhouette over a range of possible values. 

```{r}

fviz_nbclust(data, kmeans, method = "silhouette", k.max = 10) + theme_minimal() + ggtitle("Silhouette method")


```

The optimal cluster number given by the Silhouette method is 2, which is different from the value that elbow method suggests.


## 2.3 NbClust

The NbClust package provides variou indices for determining the relevant number of clusters and proposes to users the best clustering scheme from the different results by using the majority rule.


```{r}

library(NbClust)
res_nbclust = NbClust(data, min.nc = 2, max.nc = 8, method = "complete", index ="all")
plt = fviz_nbclust(res_nbclust)
plt
```

Based on the analysis above, since k = 2 has the most votes, we will choose k = 2 to be the optimal cluster number.
