
```{r ,echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(corrplot)
library(gridExtra)
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
merged <- read_csv("winequality_all.csv")
merged$red <- as.factor(merged$red )
fulldf <- merged %>% mutate(quality_group = cut(quality, breaks = c(0,4,6,10),labels=c("lo", "mid", "hi"))) 
data <- fulldf

```


## 1. Total Sulfur Dioxide

```{r, echo = FALSE, message = FALSE}
 # summary(fulldf$total.sulfur.dioxide)
fulldf %>% ggplot(aes(total.sulfur.dioxide)) + 
  geom_histogram( color = "blue",fill = "steelblue")
```

The above histogram shows the total.sulfur.dioxide has a bimodal distribution. 

## 2. Free Sulfur ioxide

```{r, echo = FALSE, message = FALSE}
#summary(fulldf$free.sulfur.dioxide)
fulldf  %>% ggplot(aes(free.sulfur.dioxide))+ 
  geom_histogram( color = "blue",fill = "steelblue")
```

The free.sulfur.dioxide has a right skewed distribution. 

```{r, echo = FALSE, message = FALSE}
ggplot(data = fulldf, aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide)) +
  geom_point() + geom_smooth()
```

By plotting free.sulfur.dioxide and total.sulfur.dioxide, we see a strong positive correlation among these two variables.

```{r, echo = FALSE, message = FALSE}
f <- ggplot(data = fulldf, aes(x = quality_group, y = free.sulfur.dioxide, color = quality_group)) +
  geom_boxplot()
t <- ggplot(data = fulldf, aes(x = quality_group, y = total.sulfur.dioxide, color = quality_group)) +
  geom_boxplot()
grid.arrange(f,t, ncol=2)
```

We discover the mid quality wine group has overall high sulfur dioxide level.